---
title: "Comparison of different model structures"
author: "Chris Terry"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

Code to generate the analysis shown in Figure 3.

```{R message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(cowplot)
library(tidybayes)
library(bayestestR)
library(ggstance)
```

# Building artificial data

The basic model has two species competing. The 'true' data is taken from a model with the form

$$ \dfrac{N_{t+1}}{ g N_{t}} = \dfrac{\lambda}{ (1+ cN_t + a C)^\theta} $$ where $N$ is the number of the focal species, $C$ is the number of the competitor species, and $a$ $b$ and $c$ are fit parameters describing the shape of the competitive response surface. The parameter 'c' is moved inside the bracket compared to the orginal to simplify interpretation. g is 'germination success' if main count is of seeds but Nt is established (germinated) plants. Poisson noise added to raw observations, but then dividing by pre-identified germination success.

```{r message=FALSE, warning=FALSE}
set.seed(3)

g=0.1
lamA = 200
interA = 0.3
intraA = 0.4
thetaA = 1.3

lamB =  300
interB = 0.8
intraB = 0.4
thetaB = 1.1


TestDensIntra<- c(0,1,5, 10,15, 20, 25, 30)
TestDensInter <- c(5, 15, 30)

### NB counts are number of competitors (i.e. excluding the focal individual!)
Intra_SpA <- data.frame(SpA = rep(TestDensIntra, each =8), SpB = 0, Type = 'INTRA_A')
Intra_SpB <- data.frame(SpA = 0, SpB =rep(TestDensIntra, each =8), Type = 'INTRA_B') 
Inter_SpA <- data.frame(SpA = 0, SpB =rep(TestDensIntra, each =8), Type = 'INTER_Asimple')
Inter_SpB <- data.frame(SpA = rep(TestDensIntra, each =8), SpB= 0, Type = 'INTER_Bsimple') 

bind_rows(Intra_SpA, Intra_SpB,  ## not including any multi-species trials - just doing the single competitors
          Inter_SpA,Inter_SpB) %>%
  mutate( SpA_Focal_Exp = lamA / ((1+ (intraA*SpA + interA*SpB))^thetaA),
          SpB_Focal_Exp = lamB / ((1+ (intraB*SpB + interB*SpA))^thetaB)) %>%
  mutate(SpA_Focal_Exp = ifelse(Type == 'INTRA_B', NA, SpA_Focal_Exp),
         SpB_Focal_Exp = ifelse(Type == 'INTRA_A', NA, SpB_Focal_Exp),
         SpB_Focal_Exp = ifelse(Type == 'INTER_Asimple', NA, SpB_Focal_Exp),
         SpA_Focal_Exp = ifelse(Type == 'INTER_Bsimple', NA, SpA_Focal_Exp)) -> DataExp

## 'True'  Value
## https://www.wolframalpha.com/input?i=1%2Fg%3D+L%2F%28%281%2BcA%29%5Eb%29%2C+solve+for+A

MonoEq_TRUE = ( ((1/(g* lamA))^(-1/thetaA)) -1 )/intraA
SpBMonoEq_TRUE = ( ((1/(g* lamB))^(-1/thetaB)) -1 )/intraB

SpA_InvFecund_TRUE  = g* lamA / ((1+ (interA*SpBMonoEq_TRUE))^thetaA)
SpB_InvFecund_TRUE  = g* lamB / ((1+ (interB*MonoEq_TRUE))^thetaB)

DataExp$SpA_Obs <- rpois(n = nrow(DataExp), lambda = DataExp$SpA_Focal_Exp)*g   ## dividing by germnation rate here
DataExp$SpB_Obs <- rpois(n = nrow(DataExp), lambda = DataExp$SpB_Focal_Exp)*g   ## dividing by germination rate here, so don't need to worry in future

MonoEq_TRUE
SpB_InvFecund_TRUE

DataforA <- DataExp %>% mutate(ObsPerFocal = SpA_Obs) %>% filter( !is.na(ObsPerFocal))
DataforB <- DataExp %>% mutate(ObsPerFocal = SpB_Obs) %>% filter( !is.na(ObsPerFocal))
```

## Plotting Raw Data

```{r}
filter(DataforA, SpB==0)%>%
  ggplot(aes(x = SpA, y = ObsPerFocal)) +
  geom_hline(yintercept = 1, linetype=2)+
  geom_vline(xintercept = MonoEq_TRUE)+
  geom_point() ->Dataplot1

filter(DataforB, SpB==0) %>%
  ggplot(aes(x = SpA, y = ObsPerFocal)) +
  geom_hline(yintercept = 1, linetype=2)+
  geom_hline(yintercept = SpB_InvFecund_TRUE)+
  geom_point() -> Dataplot2

plot_grid(Dataplot1, Dataplot2)

```

# Fitting Models

All models have two fitted parameters to simplify the comparisons. The main time sink when runnig this code is in the model compilation (not the sampling), so parrellisation is not included to aid universality.

Test models are taken from from Law and Watkinson 1981.

$$\begin{aligned}&\lambda_i-\alpha_{ii}N_i-\alpha_{ij}N_j     \\           
&\lambda_ie^{\alpha_{ii}N_i+\alpha_{ij}N_j} \\
&\lambda_ie^{-\alpha_{ii}\log(N_i+1)-\alpha_{ij}\log(N_j+1)}\\
&\lambda_i/(1+\alpha_{ii}N_i+\alpha_{ij}N_j) \\       
&\lambda_i/(1+N_i^{\alpha_{ii}}+N_j^{\alpha_{ij}}) \end{aligned}$$

```{r eval = FALSE}
# set priors
Prs <- c( set_prior("normal(25, 100)", nlpar = "lambda", lb = 0, ub = 5000),
          set_prior("normal(.1, 5)", nlpar = "p1", lb = .001, ub=2),
          set_prior("normal(.1, 5)", nlpar = "p2", lb = .001, ub=2))

# Model specifications
form_2 <- bf(ObsPerFocal~lambda-p1*SpA-p2*SpB,                   lambda+p1+p2~1,nl=TRUE)
form_3 <- bf(ObsPerFocal~lambda/exp(p1*SpA+p2*SpB),              lambda+p1+p2~1,nl=TRUE)
form_4 <- bf(ObsPerFocal~lambda/exp(p1*log(SpA+1)+p2*log(SpB+1)),lambda+p1+p2~1,nl=TRUE)
form_5 <- bf(ObsPerFocal~lambda/(1+p1*SpA+p2*SpB),               lambda+p1+p2~1,nl=TRUE)
form_6 <- bf(ObsPerFocal~lambda/(1+(SpA^p1)+(SpB^p2)),           lambda+p1+p2~1,nl=TRUE)

## Fitting for A fecundity
SpA_bayes_fit_2<- brm(form_2, data = DataforA, prior = Prs, chains=4) %>% add_criterion("loo")
SpA_bayes_fit_3<- brm(form_3, data = DataforA, prior = Prs, chains=4) %>% add_criterion("loo")
SpA_bayes_fit_4<- brm(form_4, data = DataforA, prior = Prs, chains=4) %>% add_criterion("loo")
SpA_bayes_fit_5<- brm(form_5, data = DataforA, prior = Prs, chains=4) %>% add_criterion("loo")
SpA_bayes_fit_6<- brm(form_6, data = DataforA, prior = Prs, chains=4) %>% add_criterion("loo")
## Fitting for B fecundity
SpB_bayes_fit_2<- update(SpA_bayes_fit_2, newdata = DataforB) %>% add_criterion("loo")
SpB_bayes_fit_3<- update(SpA_bayes_fit_3, newdata = DataforB) %>% add_criterion("loo")
SpB_bayes_fit_4<- update(SpA_bayes_fit_4, newdata = DataforB) %>% add_criterion("loo")
SpB_bayes_fit_5<- update(SpA_bayes_fit_5, newdata = DataforB) %>% add_criterion("loo")
SpB_bayes_fit_6<- update(SpA_bayes_fit_6, newdata = DataforB) %>% add_criterion("loo")

## Saving Fits 
save(SpA_bayes_fit_2, file = 'ModelFits/orig fits3/SpA_bayes_fit_2')
save(SpA_bayes_fit_3, file = 'ModelFits/orig fits3/SpA_bayes_fit_3')
save(SpA_bayes_fit_4, file = 'ModelFits/orig fits3/SpA_bayes_fit_4')
save(SpA_bayes_fit_5, file = 'ModelFits/orig fits3/SpA_bayes_fit_5')
save(SpA_bayes_fit_6, file = 'ModelFits/orig fits3/SpA_bayes_fit_6')
save(SpB_bayes_fit_2, file = 'ModelFits/orig fits3/SpB_bayes_fit_2')
save(SpB_bayes_fit_3, file = 'ModelFits/orig fits3/SpB_bayes_fit_3')
save(SpB_bayes_fit_4, file = 'ModelFits/orig fits3/SpB_bayes_fit_4')
save(SpB_bayes_fit_5, file = 'ModelFits/orig fits3/SpB_bayes_fit_5')
save(SpB_bayes_fit_6, file = 'ModelFits/orig fits3/SpB_bayes_fit_6')
```

### Loading Fits

```{r}
load(file = 'ModelFits/orig fits3/SpA_bayes_fit_2')
load(file = 'ModelFits/orig fits3/SpA_bayes_fit_3')
load(file = 'ModelFits/orig fits3/SpA_bayes_fit_4')
load(file = 'ModelFits/orig fits3/SpA_bayes_fit_5')
load(file = 'ModelFits/orig fits3/SpA_bayes_fit_6')
load(file = 'ModelFits/orig fits3/SpB_bayes_fit_2')
load(file = 'ModelFits/orig fits3/SpB_bayes_fit_3')
load(file = 'ModelFits/orig fits3/SpB_bayes_fit_4')
load(file = 'ModelFits/orig fits3/SpB_bayes_fit_5')
load(file = 'ModelFits/orig fits3/SpB_bayes_fit_6')


ExtractParams <- function(Model){
  Means<-summarise_draws(Model)[,1:2] 
  params <-as.list(Means$mean); names(params)<-Means$variable
  return(params)
}
```

## Extracting Values

Propogating full posterior

```{r}
Mod2_SpA <- ExtractParams(SpA_bayes_fit_2)
Mod3_SpA <- ExtractParams(SpA_bayes_fit_3)
Mod4_SpA <- ExtractParams(SpA_bayes_fit_4)
Mod5_SpA <- ExtractParams(SpA_bayes_fit_5)
Mod6_SpA <- ExtractParams(SpA_bayes_fit_6)


Mod2_SpA <- SpA_bayes_fit_2 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept ) 
Mod3_SpA <- SpA_bayes_fit_3 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept ) 
Mod4_SpA <- SpA_bayes_fit_4 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept ) 
Mod5_SpA <- SpA_bayes_fit_5 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept ) 
Mod6_SpA <- SpA_bayes_fit_6 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept )

Mod2_SpB <- ExtractParams(SpB_bayes_fit_2)
Mod3_SpB <- ExtractParams(SpB_bayes_fit_3)
Mod4_SpB <- ExtractParams(SpB_bayes_fit_4)
Mod5_SpB <- ExtractParams(SpB_bayes_fit_5)
Mod6_SpB <- ExtractParams(SpB_bayes_fit_6)

Mod2_SpB <- SpB_bayes_fit_2 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept ) 
Mod3_SpB <- SpB_bayes_fit_3 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept ) 
Mod4_SpB <- SpB_bayes_fit_4 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept ) 
Mod5_SpB <- SpB_bayes_fit_5 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept ) 
Mod6_SpB <- SpB_bayes_fit_6 %>% spread_draws(b_lambda_Intercept,b_p1_Intercept,b_p2_Intercept )

```

## Calculating Monoculture equilibria and invasion growth rates

```{r}
## Model 2  ObsPerFocal~lambda-p1*SpA-p2*SpB,    
## (lambda-1) /(a)
MonoEq_2 = (Mod2_SpA$b_lambda_Intercept-1) /  Mod2_SpA$b_p1_Intercept 

#  Model 3  ObsPerFocal~lambda*exp(-p1*SpA-p2*SpB), 
##  MonoEq_3 = log(lambda)/a
MonoEq_3 = log(Mod3_SpA$b_lambda_Intercept) /  Mod3_SpA$b_p1_Intercept 

## model 4 ObsPerFocal~lambda*exp(-p1*log(SpA+1)-p2*log(SpB+1))
## MonoEq_4 =  lambda^(1/a)
MonoEq_4 = Mod4_SpA$b_lambda_Intercept ^ (1/Mod4_SpA$b_p1_Intercept)

## Model 5  ObsPerFocal~lambda/(1+p1*SpA+p2*SpB),               
## MonoEq_5 = (lambda-1)/a
MonoEq_5 = (Mod5_SpA$b_lambda_Intercept-1)/Mod5_SpA$b_p1_Intercept

## Model 6 ObsPerFocal~lambda/(1+(SpA^p1)+(SpB^p2)),           
## MonoEq_6 = (lambda-1)^(1/a)  ## 
MonoEq_6 = (Mod6_SpA$b_lambda_Intercept-1)^(1/Mod6_SpA$b_p1_Intercept)  ## 

SpB_InvFecund2 <- Mod2_SpB$b_lambda_Intercept-Mod2_SpB$b_p1_Intercept*MonoEq_2               
SpB_InvFecund3 <- Mod3_SpB$b_lambda_Intercept/exp(Mod3_SpB$b_p1_Intercept*MonoEq_3)        
SpB_InvFecund4 <- Mod4_SpB$b_lambda_Intercept/exp(Mod4_SpB$b_p1_Intercept*log(MonoEq_4+1))
SpB_InvFecund5 <- Mod5_SpB$b_lambda_Intercept/(1+Mod5_SpB$b_p1_Intercept*MonoEq_5)              
SpB_InvFecund6 <- Mod6_SpB$b_lambda_Intercept/(1+(MonoEq_6^Mod6_SpB$b_p1_Intercept))          
```

## Finding mean and 90% interval of IGRs

```{r message=FALSE, warning=FALSE}
data.frame(SpA =  seq(0,40, by = 0.1),
           SpB = 0) %>%  
  mutate(ID = 1:n()) %>%
  group_by(SpA, ID) %>%
  add_epred_draws(SpA_bayes_fit_2, value = 'EPred_2') %>% summarise( Mod_2_LV  = mean(EPred_2)) %>%
  add_epred_draws(SpA_bayes_fit_3, value = 'EPred_3') %>% summarise( Mod_3_Exp = mean(EPred_3)) %>%
  add_epred_draws(SpA_bayes_fit_4, value = 'EPred_4') %>% summarise( Mod_4_EBH = mean(EPred_4)) %>%
  add_epred_draws(SpA_bayes_fit_5, value = 'EPred_5') %>% summarise( Mod_5_BH  = mean(EPred_5)) %>%
  add_epred_draws(SpA_bayes_fit_6, value = 'EPred_6') %>% summarise( Mod_6_Exp = mean(EPred_6))  -> Mean_Epred_allMods

data.frame(SpA =  seq(0,40, by = 0.1),SpB = 0) %>%
  mutate( ExpSeedsPerFocal = lamA / ((1+ (intraA*SpA + interA*SpB))^thetaA),
          ExpAdultsPerFocal = g * ExpSeedsPerFocal) -> DataExp_Across

## Response of B to increasing competitors A

data.frame(SpA = seq(0,40, by = 0.1),
           SpB = 0) %>%  
  mutate(ID = 1:n()) %>%
  group_by(SpA, ID) %>%
  add_epred_draws(SpB_bayes_fit_2, value = 'EPred_2') %>% summarise( Mod_2_LV  = mean(EPred_2)) %>%
  add_epred_draws(SpB_bayes_fit_3, value = 'EPred_3') %>% summarise( Mod_3_Exp = mean(EPred_3)) %>%
  add_epred_draws(SpB_bayes_fit_4, value = 'EPred_4') %>% summarise( Mod_4_EBH = mean(EPred_4)) %>%
  add_epred_draws(SpB_bayes_fit_5, value = 'EPred_5') %>% summarise( Mod_5_BH  = mean(EPred_5)) %>%
  add_epred_draws(SpB_bayes_fit_6, value = 'EPred_6') %>% summarise( Mod_6_Exp = mean(EPred_6)) -> Mean_Epred_allMods_BINVASION


data.frame(SpA = seq(0,40, by = 0.1),SpB = 0) %>%
  mutate( SpB_ExpSeedsPerFocal= lamB / ((1+ (intraB*SpB + interB*SpA))^thetaB),
          ExpAdultsPerFocal = g * SpB_ExpSeedsPerFocal) -> DataExp_B_Invasion


data.frame(Model = c(rep(c('Mod_2_LV'  ,'Mod_3_Exp' ,'Mod_4_EBH', 'Mod_5_BH'  ,
                           'Mod_6_Exp'), each = 4000),  'MOD_TRUE'),
           MonoEq = c( MonoEq_2, MonoEq_3, MonoEq_4, MonoEq_5, MonoEq_6, MonoEq_TRUE),
           InvasionFecund = c(SpB_InvFecund2,SpB_InvFecund3,
                              SpB_InvFecund4,SpB_InvFecund5,
                              SpB_InvFecund6, SpB_InvFecund_TRUE)) -> fullpost

HDI2eq <- fullpost %>% filter(Model == "Mod_2_LV") %>% bayestestR::hdi()
HDI3eq <- fullpost %>% filter(Model == "Mod_3_Exp") %>% bayestestR::hdi()
HDI4eq <- fullpost %>% filter(Model == "Mod_4_EBH") %>% bayestestR::hdi()
HDI5eq <- fullpost %>% filter(Model == "Mod_5_BH") %>% bayestestR::hdi()
HDI6eq <- fullpost %>% filter(Model == "Mod_6_Exp") %>% bayestestR::hdi()

mean2eq <- fullpost %>% filter(Model == "Mod_2_LV") %>% select(-1) %>% colMeans(na.rm = TRUE)
mean3eq <- fullpost %>% filter(Model == "Mod_3_Exp")  %>% select(-1) %>% colMeans(na.rm = TRUE)
mean4eq <- fullpost %>% filter(Model == "Mod_4_EBH")  %>% select(-1) %>% colMeans(na.rm = TRUE)
mean5eq <- fullpost %>% filter(Model == "Mod_5_BH")  %>% select(-1) %>% colMeans(na.rm = TRUE)
mean6eq <- fullpost %>% filter(Model == "Mod_6_Exp")  %>% select(-1) %>% colMeans(na.rm = TRUE)

fullpost %>%
  group_by(Model)%>%
  summarise( MeanMonoEq = mean(MonoEq ), 
             MeanInvasionFecund = mean(InvasionFecund))

Model = c('Mod_2_LV'  ,'Mod_3_Exp' ,'Mod_4_EBH', 'Mod_5_BH', 'Mod_6_Exp')
CIhi_eq = c(HDI2eq$CI_high[1], HDI3eq$CI_high[1],
            HDI4eq$CI_high[1],HDI5eq$CI_high[1], HDI6eq$CI_high[1])
CIlo_eq = c(HDI2eq$CI_low[1], HDI3eq$CI_low[1],
            HDI4eq$CI_low[1],  HDI5eq$CI_low[1],HDI6eq$CI_low[1])
ypos_eq = c(-3, -2.5, -2, -1.5, -1)/2
xval_eq = c(mean2eq[1], mean3eq[1],
            mean4eq[1], mean5eq[1],  mean6eq[1])
segdat_equil = data.frame(Model, CIlo_eq, CIhi_eq, ypos_eq, xval_eq)

CIhi_inv = c(HDI2eq$CI_high[2],  HDI3eq$CI_high[2],
             HDI4eq$CI_high[2], HDI5eq$CI_high[2], HDI6eq$CI_high[2])
CIlo_inv = c(HDI2eq$CI_low[2], HDI3eq$CI_low[2],
             HDI4eq$CI_low[2],  HDI5eq$CI_low[2], HDI6eq$CI_low[2])
xpos_inv = c(0,1,2,3,4)+34
yval_inv = c(mean2eq[2],  mean3eq[2],
             mean4eq[2], mean5eq[2], mean6eq[2])
segdat_inv = data.frame(Model, CIlo_inv, CIhi_inv, xpos_inv, yval_inv)

```

# Plots

## Plot A

```{r}

Colours <- c( 'Mod_2_LV'='#648FFF' , 'Mod_3_Exp' = '#785EF0' ,
              'Mod_4_EBH' = '#DC267F','Mod_5_BH' ='#FE6100'  ,
              'Mod_6_Exp'= '#FFB000',  'MOD_TRUE' ='black')# IBM design library colours

Mean_Epred_allMods %>%
  pivot_longer(cols = starts_with('Mod_'),
               names_to = 'Model', 
               values_to = 'Prediction') %>%
  mutate(Prediction = ifelse(Prediction <0, 0.001,  Prediction) ) %>%
  ggplot(aes(x = SpA)) +
   geom_rect(aes( xmin = -10, xmax = 50, ymax = 1, ymin =0), fill = 'grey90')+
  geom_line(aes(y = Prediction, col = Model))+
  geom_line(aes(y = ExpAdultsPerFocal), data = DataExp_Across, linewidth =1)+
  geom_point(aes(y = ObsPerFocal), data = filter(DataforA, SpB==0),
             shape = 1, size = 3, alpha = 0.75) +
  geom_point(x=MonoEq_TRUE, y = 1,  size = 3, alpha = 0.75, shape = 4) +
  theme_minimal()+
  theme(legend.position = 'bottom', 
        panel.grid  =  element_blank())+
  scale_color_manual(values = Colours)+
  scale_fill_manual(values = Colours)+
  geom_line(aes(y = ExpAdultsPerFocal), data = DataExp_Across, size =1)+
  #scale_y_log10(limits = c(0.5, 30))+
  coord_cartesian(ylim = c(-1.5, 8), xlim = c(0, 50))+
  scale_y_continuous(breaks = c(0,1:8)) +
  ylab('Species A Reproductive Rate')+
  xlab('Number of Intraspecific Competitors')+
  guides(shape = 'none', size = 'none', col = 'none')+
  geom_pointrangeh(data = segdat_equil,
                   aes(x = xval_eq, y = ypos_eq,
                       xmin = CIlo_eq, xmax = CIhi_eq,
                       color = Model), inherit.aes = F) +
  geom_text(x = MonoEq_TRUE, y =5,
            label = 'True Sp. A\nEquilibrium' , col = "grey60", 
            angle = 90, size =3, fontface='plain')+
  geom_vline(xintercept = MonoEq_TRUE, lty = 2, color = "grey60") +
  theme(legend.justification = 'center',
        legend.position = c(0.6, 0.8),
        legend.title = element_text(size = 16),  
        legend.text = element_text(size = 16), 
        legend.background = element_rect(colour = 'white'))+
  ggtitle('a) Species A monoculture equilibrium fit') -> SpeciesAEqPlot

```

## Plot B

```{r}
Mean_Epred_allMods_BINVASION %>%
  pivot_longer(cols = starts_with('Mod_'),names_to = 'Model', 
               values_to = 'Prediction') %>%
  filter( SpA<31) %>%
  ggplot(aes(x = SpA)) +
  geom_rect(aes( xmin = -10, xmax = 50, ymax = 1, ymin =0), fill = 'grey90')+
  geom_line(aes(y = Prediction, col = Model))+
  geom_line(aes(y = ExpAdultsPerFocal),
            data = filter(DataExp_B_Invasion, SpA <31), linewidth =1)+
  geom_point(aes(y = ObsPerFocal),
             data = filter(DataforB, SpB==0),
             shape =1, size = 3, alpha = 0.75) +
  scale_color_manual(values = Colours)+
  scale_fill_manual(values = Colours)+
  theme_minimal()+
  theme(legend.position = 'bottom', 
        panel.grid =  element_blank())+
  ylab('Species B Reproductive Rate')+
  xlab('Number of Interspecific Competitors')+
  coord_cartesian(ylim = c(0, 5), xlim = c(0, 40))+
  scale_x_continuous(breaks = c(0,10, 20, 30)) +
  geom_hline(yintercept = SpB_InvFecund_TRUE, lty = 2, color = "grey60") +
  geom_text(x = 5, y =SpB_InvFecund_TRUE, label = 'True Invasion\nGrowth Rate' ,
            col = "grey60", size =3, fontface='plain')+
  geom_vline(xintercept = MonoEq_TRUE, lty = 2, color = "grey60") +
  geom_text(x = MonoEq_TRUE, y =4,
            label = 'True Sp. A\nEquilibrium' , col = "grey60",
            angle = 90, size =3, fontface='plain')+
  geom_point(x=MonoEq_TRUE, y = SpB_InvFecund_TRUE,  size = 3, alpha = 0.75, shape = 4) +
  geom_pointrange(data = segdat_inv,
                  aes(y = yval_inv, x = xpos_inv, ymin = CIlo_inv,
                      ymax = CIhi_inv, color = Model), inherit.aes = F) +
  ggtitle('b) Species B invasion growth rate') -> SpeciesB_InvPlot

```

## Combining

```{R fig.height=4, fig.width=8}

model_plots = plot_grid(SpeciesAEqPlot,
                        SpeciesB_InvPlot+guides(fill=FALSE,col=FALSE),
                        ncol = 2, rel_heights = c(3,3))

model_plots

ggsave(plot = model_plots,'Figures/Figure3.pdf', width = 8, height = 4)
ggsave(plot = model_plots,'Figures/Figure3.png', width = 8, height = 4,
       dpi= 600, bg = 'white')

```

# Model Comparisons

```{r}
loo_compare( list(m2 = SpA_bayes_fit_2$criteria$loo,
                  m3 = SpA_bayes_fit_3$criteria$loo,
                  m4 = SpA_bayes_fit_4$criteria$loo,
                  m5 = SpA_bayes_fit_5$criteria$loo,
                  m6 = SpA_bayes_fit_6$criteria$loo)) %>%
  as.data.frame %>% 
  rownames_to_column('model') %>%
  arrange(model)  %>%
  select( model,partA_elpd_diff = elpd_diff)-> partA_ELPD

loo_compare( list(m2 = SpB_bayes_fit_2$criteria$loo,
                  m3 = SpB_bayes_fit_3$criteria$loo,
                  m4 = SpB_bayes_fit_4$criteria$loo,
                  m5 = SpB_bayes_fit_5$criteria$loo,
                  m6 = SpB_bayes_fit_6$criteria$loo)) %>%
  as.data.frame %>% 
  rownames_to_column('model') %>%
  arrange(model) %>%
  select( model,partB_elpd_diff =elpd_diff) -> partB_ELPD

BothELPD <- left_join(partA_ELPD, partB_ELPD)

BothELPD$Model <- c('Mod_2_LV'  ,'Mod_3_Exp' ,'Mod_4_EBH', 'Mod_5_BH'  ,
                           'Mod_6_Exp')

BothELPD$IGR = yval_inv

BothELPD %>%
  knitr::kable( digits = 3)

SpB_InvFecund_TRUE

```

# Session Info
```{r}
sessionInfo()
```