---
title: "Comparison of Error Structures"
author: "Chris Terry"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

This script includes the code to generate the example in Figure 4. 

```{R message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(cowplot)
library(tidybayes)
library(posterior)
```

# Generating simulated data

Assuming true model is Beverton-Holt with log-normal variation structure.
Generating noise with draws with parameters adjusted to maintain mean expected value.+
A high number of replicates are used to try and minimise fitting noise, but exact results are still highly random seed dependent.

```{r}
set.seed(4)  

### True Values
lam_1 = 50
lam_2 = 30
a11 = 1.1
a22 = 1
a21 = 1.5
a12  = 1.25

log_sigma1 = 1.3
log_sigma2 = 0.6

True_values = data.frame(variable = c('b_lambda_FocalSp1','b_lambda_FocalSp2',  
                                      'b_intra_FocalSp1' ,'b_intra_FocalSp2' ,  
                                      'b_inter_FocalSp1' ,'b_inter_FocalSp2'),
                         TRUE_value = c(lam_1, lam_2, a11, a22, a12, a21))

### Generating Fake Data
reps = 100  
Levels <- c(2,5,12)

## Sp 1  as focal
Sp1_intra <- expand_grid(Focal = 'Sp1', N_focal = c(0,Levels),
                         N_comp = 0, rep = 1:reps, Test = 'INTRA')
Sp1_inter <- expand_grid(Focal = 'Sp1',N_focal = 0, N_comp = Levels,
                         rep = 1:reps, Test = 'INTER' )

bind_rows(Sp1_intra,Sp1_inter ) %>%
  mutate(ExpectedSeeds =  lam_1 / (1+ a11*N_focal + a12*N_comp  ),
         ExpectedLogSeeds  = log(ExpectedSeeds),
         AdjustLogExpect = ExpectedLogSeeds - (log_sigma1^2)/2 ) %>%
  rowwise() %>%
  mutate( ObsSeeds = round(rlnorm( n = 1,
                                   meanlog = AdjustLogExpect, 
                                   sdlog = log_sigma1),
                           0)) -> Sp1_Obs
Sp1_Obs %>%
  filter( N_focal+N_comp == 0) %>%
  ungroup()%>%
  summarise(mean(ObsSeeds))

### Species 2 as focal
Sp2_intra <- expand_grid(Focal = 'Sp2', N_focal = Levels,
                         N_comp = 0,rep = 1:reps, Test = 'INTRA')
Sp2_inter <- expand_grid(Focal = 'Sp2',N_focal = 0,
                         N_comp =Levels, rep = 1:reps,  Test = 'INTER' )

bind_rows(Sp2_intra,Sp2_inter ) %>%
  mutate(ExpectedSeeds =  lam_2 / (1+ a22*N_focal + a21*N_comp  ),
         ExpectedLogSeeds  = log(ExpectedSeeds),
         AdjustLogExpect = ExpectedLogSeeds - (log_sigma2^2)/2 ) %>%
  rowwise() %>%
  mutate( ObsSeeds = round(rlnorm( n = 1,
                                   meanlog = AdjustLogExpect,
                                   sdlog = log_sigma2),
                           0)) -> Sp2_Obs

bind_rows(Sp1_Obs, Sp2_Obs)%>%
  mutate( LogObsSeeds = log(ObsSeeds+0.01))%>% ## to prevent mathematical errors
  mutate(Competitors = N_focal+N_comp) -> BothData

```

# Plotting Raw Data

```{r}
BothData%>%
  mutate(Competitors = N_focal+N_comp) %>%
  mutate(Competition= recode(Test,
                             'INTRA' = 'Intraspecific',
                             'INTER' = 'Interspecific'),
         Focal = recode(Focal ,
                             'Sp1' = 'Species 1',
                             'Sp2' = 'Species 2')) %>%
  ggplot(aes( x =Competitors ))+
  geom_point(aes(y=ObsSeeds), size = 0.8, alpha = 0.1)+
  geom_point(aes(y=ExpectedSeeds))+
  geom_line(aes(y=ExpectedSeeds))+
  facet_wrap(Focal ~Competition, 
             scales = 'free_y',
             labeller = label_both, nrow =1)+
  scale_y_log10()+
  theme_minimal()+
  xlab('Number of Competitors')+
  ylab('Observed Focal Species\nOutput per Individual')+
  scale_x_continuous(breaks = c(0,Levels)) -> A_RawData
A_RawData
```

# Model Fitting

```{r eval = FALSE}
m1_prior <- prior_string("uniform(0,50)",lb=0.1,ub=50,nlpar="lambda") +
  prior_string("uniform(0,5)",lb=0,ub=5,nlpar="intra") +
  prior_string("uniform(0,5)",lb=0,ub=5,nlpar="inter")

## Log Scale fitting
fit_m1_log_splitsigma <- brm(brmsformula(formula = LogObsSeeds ~ log( lambda / (1 + intra * N_focal  + inter * N_comp)),
                                         nonlinear = list(lambda ~ 0 + Focal, intra ~ 0 + Focal,
                                                          inter ~  0 + Focal, sigma ~0+ Focal),
                                         nl = TRUE),
                             data = BothData, prior = m1_prior,
                             cores = getOption("mc.cores", 4))

## Natural Scale Gaussian Fitting
fit_m1_direct_splitsigma <- brm(brmsformula(formula = ObsSeeds ~ lambda / (1 + intra * N_focal  + inter * N_comp),
                                            nonlinear = list(lambda ~ 0 + Focal, intra ~ 0 + Focal,
                                                             inter ~  0 + Focal,sigma ~0+ Focal ),
                                            nl = TRUE),
                                data = BothData, prior = m1_prior,
                                cores = getOption("mc.cores", 4))

## Negative Binomial fitting
fit_m1_direct_NegBimom <- brm(brmsformula(formula = ObsSeeds ~ lambda / (1 + intra * N_focal  + inter * N_comp),
                                          family = negbinomial(link = 'identity'), 
                                          nonlinear = list(lambda ~ 0 + Focal, intra ~ 0 + Focal,
                                                           inter ~  0 + Focal, shape~0+Focal),
                                          nl = TRUE),
                              data = BothData, prior = m1_prior,
                              cores = getOption("mc.cores", 4))

## Poisson fitting
fit_m1_direct_Poiss <- brm(brmsformula(formula = ObsSeeds ~ lambda / (1 + intra * N_focal  + inter * N_comp),
                                       family = poisson(link = 'identity'), 
                                       nonlinear = list(lambda ~ 0 + Focal, intra ~ 0 + Focal,
                                                        inter ~  0 + Focal),
                                       nl = TRUE),
                           data = BothData, prior = m1_prior,
                           cores = getOption("mc.cores", 4))

save(fit_m1_log_splitsigma, file='ModelFits/fit_m1_log_splitsigma')
save(fit_m1_direct_splitsigma, file='ModelFits/fit_m1_direct_splitsigma')
save(fit_m1_direct_NegBimom, file='ModelFits/fit_m1_direct_NegBimom')
save(fit_m1_direct_Poiss, file='ModelFits/fit_m1_direct_Poiss')
```



```{r}
load(file='ModelFits/fit_m1_log_splitsigma')
load(file='ModelFits/fit_m1_direct_splitsigma')
load(file='ModelFits/fit_m1_direct_NegBimom')
load(file='ModelFits/fit_m1_direct_Poiss')

```

### Best Estimates of parameters

```{r}
as_draws_array(fit_m1_direct_splitsigma) %>%
  summarise_draws(.x = .,
                  "mean","median","sd","mad","quantile2",~quantile2(.x, probs = c(0.20, 0.80)))%>%
  mutate( Model = 'Gaussian')-> Gaussian_params

as_draws_array(fit_m1_log_splitsigma) %>%
  summarise_draws(.x = .,
                  "mean","median","sd","mad","quantile2",~quantile2(.x, probs = c(0.20, 0.80)))%>%
  mutate( Model = 'Log transform,\nthen Gaussian')-> LogGaussian_params

as_draws_array(fit_m1_direct_Poiss) %>%
  summarise_draws(.x = .,
                  "mean","median","sd","mad","quantile2",~quantile2(.x, probs = c(0.20, 0.80)))%>%
  mutate( Model = 'Poisson') -> Pois_params

as_draws_array(fit_m1_direct_NegBimom) %>%
  summarise_draws(.x = .,
                  "mean","median","sd","mad","quantile2",~quantile2(.x, probs = c(0.20, 0.80))) %>%
  filter( variable !='b_shape_FocalSp1')%>%
  filter( variable !='b_shape_FocalSp2')%>%
  mutate( Model = 'Negative-Binomial') -> NegBinom_params

bind_rows( Gaussian_params,LogGaussian_params,
           Pois_params,NegBinom_params ) %>%
  filter( str_starts(variable, 'b_' )) %>%
  filter( variable !='b_sigma_FocalSp1')%>%
  filter( variable !='b_sigma_FocalSp2')%>%
  left_join( True_values, by = "variable") %>%
  separate(remove=FALSE, variable,
           into = c('b', 'VAR', 'Focal' ))%>%
  mutate( Focal = str_sub(Focal, 6, 8))-> Param_df

Param_df%>%
  mutate(VAR= recode(VAR,
                     'inter' = ' α Intraspecific',
                     'intra' = 'α Interspecific',
                     'lambda'= 'λ'),
         Focal = recode(Focal ,
                             'Sp1' = 'Species 1',
                             'Sp2' = 'Species 2')) %>%
  ggplot(aes( y =Model , x = mean))+
  geom_segment(aes( x= q5, xend = q95, yend = Model , col =Model ), linewidth = 1)+
  geom_segment(aes( x= q20, xend = q80, yend = Model, col =Model  ), linewidth = 2)+
  geom_point(size =2)+
  # geom_errorbarh(aes( xmin = q5, xmax = q95))+
  facet_grid(Focal~VAR, 
             scales = 'free_x')+
  geom_vline(aes( xintercept = TRUE_value))+
  scale_y_discrete( breaks = c( 'Gaussian',
                                'Log transform,\nthen Gaussian',
                                'Poisson',
                                'Negative-Binomial')) +
  xlab('Parameter Value')+
  ylab('Transformation\nand Error Approach')+
  theme_minimal()+
  guides(col = 'none')+
  scale_colour_brewer(type = 'qual', palette = 'Dark2') -> B_Param_estimates
B_Param_estimates
```

### Consequences for Predicted Coexistence

```{R}
## for plotting
niche_differentiation <- seq(from = -.25, to = 1, by = 0.001)
niche_overlap <- 1-niche_differentiation
fitness_ratio_min <- niche_overlap
fitness_ratio_max <- 1/niche_overlap
coexistarea_df <- data.frame(niche_diff = niche_differentiation,
                             min_fitness_ratio = fitness_ratio_min,
                             max_fitness_ratio = fitness_ratio_max)

## Log scale  ### Full posterior
LOG_post <- as_draws_df(fit_m1_log_splitsigma)
ND_logged = 1 - sqrt((LOG_post$b_intra_FocalSp1  *LOG_post$b_intra_FocalSp2) / (LOG_post$b_inter_FocalSp1 * LOG_post$b_inter_FocalSp2)    )
RFD_logged =  ((LOG_post$b_lambda_FocalSp1-1) / (LOG_post$b_lambda_FocalSp2-1)) *  sqrt((LOG_post$b_inter_FocalSp2*LOG_post$b_intra_FocalSp2) / (LOG_post$b_inter_FocalSp1 * LOG_post$b_intra_FocalSp1  )    )

### Natural Scale # Full posterior
Ns <- as_draws_df(fit_m1_direct_splitsigma)
ND_natural = 1 - sqrt((Ns$b_intra_FocalSp1  *Ns$b_intra_FocalSp2) / (Ns$b_inter_FocalSp1 * Ns$b_inter_FocalSp2)    )
RFD_natural =  ((Ns$b_lambda_FocalSp1-1) / (Ns$b_lambda_FocalSp2-1)) *  sqrt((Ns$b_inter_FocalSp2*Ns$b_intra_FocalSp2) / (Ns$b_inter_FocalSp1 * Ns$b_intra_FocalSp1  )    )

### Negative Binomial # Full posterior
NBs <- as_draws_df(fit_m1_direct_NegBimom)
ND_NegBimom = 1 - sqrt((NBs$b_intra_FocalSp1  *NBs$b_intra_FocalSp2) / (NBs$b_inter_FocalSp1 * NBs$b_inter_FocalSp2)    )
RFD_NegBimom =  ((NBs$b_lambda_FocalSp1-1) / (NBs$b_lambda_FocalSp2-1)) *  sqrt((NBs$b_inter_FocalSp2*NBs$b_intra_FocalSp2) / (NBs$b_inter_FocalSp1 * NBs$b_intra_FocalSp1  )    )

### Pois # Full posterior
POI <- as_draws_df(fit_m1_direct_Poiss)
ND_Poiss = 1 - sqrt((POI$b_intra_FocalSp1  *POI$b_intra_FocalSp2) / (POI$b_inter_FocalSp1 * POI$b_inter_FocalSp2)    )
RFD_Poiss =  ((POI$b_lambda_FocalSp1-1) / (POI$b_lambda_FocalSp2-1)) *  sqrt((POI$b_inter_FocalSp2*POI$b_intra_FocalSp2) / (POI$b_inter_FocalSp1 * POI$b_intra_FocalSp1  )    )



## 'True' Values
ND_true = 1 - sqrt((a11 *a22) / (a12 * a21)    )
RFD_true =  ((lam_1-1)/ (lam_2-1)) *  sqrt((a21*a22) / (a12* a11 )    )


CoexistPlane <-data.frame(X = c( 'Gaussian',
                                 'Log transform,\nthen Gaussian',
                                 'Poisson',
                                 'Negative-Binomial','True' ),
                          NicheDiffsMEAN = c(  mean(ND_natural), 
                                               mean(ND_logged), 
                                               mean(ND_Poiss),
                                               mean(ND_NegBimom), 
                                               ND_true),
                          RelFitDiffMEAN = c( mean(RFD_natural),
                                              mean(RFD_logged),
                                              mean(RFD_Poiss),
                                              mean(RFD_NegBimom),
                                              RFD_true),
                          NicheDiffsLOW = c( quantile2(ND_natural, 0.05),
                                             quantile2(ND_logged, 0.05), 
                                             quantile2(ND_Poiss, 0.05), 
                                             quantile2(ND_NegBimom, 0.05), 
                                             ND_true),
                          RelFitDiffLOW = c(quantile2(RFD_natural, 0.05),
                                            quantile2(RFD_logged, 0.05),
                                            quantile2(RFD_Poiss, 0.05),
                                            quantile2(RFD_NegBimom, 0.05),
                                            RFD_true),
                          NicheDiffsHIGH = c( quantile2(ND_natural, 0.95),
                                              quantile2(ND_logged, 0.95), 
                                              quantile2(ND_Poiss, 0.95), 
                                              quantile2(ND_NegBimom, 0.95), 
                                              ND_true),
                          RelFitDiffHIGH = c(quantile2(RFD_natural, 0.95),
                                             quantile2(RFD_logged, 0.95),
                                             quantile2(RFD_Poiss, 0.95),
                                             quantile2(RFD_NegBimom, 0.95),
                                             RFD_true))


CoexistPlane


FourCols <-RColorBrewer::brewer.pal(4, 'Dark2')

CoexistPlane %>%
  ggplot()+
  geom_ribbon(data = coexistarea_df, aes(x = niche_diff, ymin = min_fitness_ratio,
                                         ymax = max_fitness_ratio), fill = 'grey80')+
  geom_line(data = coexistarea_df, aes(x = niche_diff, y = max_fitness_ratio)) +
  geom_line(data = coexistarea_df,  aes(x = niche_diff, y = min_fitness_ratio)) +
  geom_point( aes( x = NicheDiffsMEAN, y = RelFitDiffMEAN, col = X), size = 4)+
  geom_errorbar( aes( x = NicheDiffsMEAN, ymin = RelFitDiffLOW,
                      ymax = RelFitDiffHIGH, col = X))+
  geom_errorbarh( aes( xmin = NicheDiffsLOW, xmax = NicheDiffsHIGH,
                       y = RelFitDiffMEAN, col = X))+
  scale_y_log10()+
  coord_cartesian(xlim = c(-0.05,0.5), ylim = c(0.5, 2) )+
  scale_colour_manual(name = 'Model', values = c(FourCols, 'black' ))+
  theme_minimal()+
  ylab('Fitness Ratio\n(Sp 1 : Sp 2)' )+
  xlab('Niche Difference') -> C_CorexistPlane
C_CorexistPlane
```

# Combining  parts of the figure together

```{r message=FALSE, warning=FALSE}
Bottom<- plot_grid(B_Param_estimates,  C_CorexistPlane, labels = c('b)', 'c)'))
plot_grid(A_RawData,Bottom, nrow = 2, labels = c('a)', ''))
ggsave('Figures/Fig4_ErrorStruct.png', width = 10, height = 6, dpi = 600, bg = 'white')
```


# Plotting model fits to data



```{r}

Ls = as.list(colMeans(LOG_post))
Ns = as.list(colMeans(Ns))
NBs = as.list(colMeans(NBs)  )
POI = as.list(colMeans(POI))


expand.grid(Focal = c('Sp1', 'Sp2'), 
            N_focal = 0:15,
            N_comp = 0:15  )%>%
  filter( N_focal ==0 | N_comp == 0) %>%
  mutate( Test = ifelse( N_comp == 0, 'INTRA', 'INTER')) %>%
  mutate( `Log transform,\nthen Gaussian` = ifelse(Focal =='Sp1',
                                                   Ls$b_lambda_FocalSp1 / (1 + Ls$b_intra_FocalSp1 * N_focal  + Ls$b_inter_FocalSp1 * N_comp),
                                                   Ls$b_lambda_FocalSp2 / (1 + Ls$b_intra_FocalSp2 * N_focal  + Ls$b_inter_FocalSp2 * N_comp)),
          Gaussian = ifelse(Focal =='Sp1',
                            Ns$b_lambda_FocalSp1 / (1 + Ns$b_intra_FocalSp1 * N_focal  + Ns$b_inter_FocalSp1 * N_comp),
                            Ns$b_lambda_FocalSp2 / (1 + Ns$b_intra_FocalSp2 * N_focal  + Ns$b_inter_FocalSp2 * N_comp)),
          `Negative-Binomial` = ifelse(Focal =='Sp1',
                                       NBs$b_lambda_FocalSp1 / (1 + NBs$b_intra_FocalSp1 * N_focal  + NBs$b_inter_FocalSp1 * N_comp),
                                       NBs$b_lambda_FocalSp2 / (1 + NBs$b_intra_FocalSp2 * N_focal  + NBs$b_inter_FocalSp2 * N_comp)),
          Poisson = ifelse(Focal =='Sp1',
                           POI$b_lambda_FocalSp1 / (1 + POI$b_intra_FocalSp1 * N_focal  + POI$b_inter_FocalSp1 * N_comp),
                           POI$b_lambda_FocalSp2 / (1 + POI$b_intra_FocalSp2 * N_focal  + POI$b_inter_FocalSp2 * N_comp))) -> Predictions


Predictions %>%
  pivot_longer(names_to = 'Model',
               values_to = 'Prediction',
               cols = 5:8)%>%
  mutate(Competitors = N_focal+N_comp) %>%
  mutate(Competition= recode(Test,
                             'INTRA' = 'Intraspecific',
                             'INTER' = 'Interspecific')) -> PredictionsLong

BothData%>%
  mutate(Competitors = N_focal+N_comp) %>%
  mutate(Competition= recode(Test,
                             'INTRA' = 'Intraspecific',
                             'INTER' = 'Interspecific')) %>%
  ggplot(aes( x =Competitors ))+
  geom_point(aes(y=ObsSeeds), alpha = 0.2, shape = 1)+
  geom_point(aes(y=ExpectedSeeds))+
  geom_line(aes(y=ExpectedSeeds))+
  geom_line(aes(y=Prediction, col = Model), data = PredictionsLong)+
  facet_grid(Focal ~Competition, 
             scales = 'free_y',
             labeller = label_both)+
  scale_colour_manual(name = 'Model', values =FourCols)+
  theme_minimal()+
  theme(legend.position = 'bottom')+
  xlab('Number of Competitors')+
  ylab('Observed Focal Species Output\nper individual')+
  scale_x_continuous(breaks = c(0,Levels))+
  scale_y_log10()-> SI2

ggsave( 'Figures/FigureSI3.png',SI2,  height= 6, width=6, dpi = 600, bg = 'white')

```


# Session Info
```{r}
sessionInfo()
```
